---
title: "DQF Transforms"
output: html_document
date: "2023-06-12"
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval=FALSE, include=FALSE,results='hide'}
source("00. datasets.R")
source("01. dqf-outlier.R")
source("02. dqf-subset.R")
source("03. extract-dqf-subset.R")
```

# Export
```{r, eval=FALSE, include=FALSE,results='hide'}
knitr::purl("10. dqf-transforms.Rmd")
```

### Scale functions (function specific)

#### `scale.dqf.max`: Given dqfs, where rows and columns are functions' x and y values, scale each function to have min=0, max=1.

```{r}
scale.dqf.max <- function(dqf){
  n.functions <- length(dqf[,1])
  ret <- dqf
  
  for(i in 1:n.functions){
    func.max <- max(dqf[i,])
    for(j in 1:length(dqf[i,])){
      ret[i,j] <- dqf[i,j]/func.max
    }
  }
  
  return(ret)
}
```

#### `scale.dqf.sum`: Given dqfs, where rows and columns are functions' x and y values, scale each function to its sum.

```{r}
scale.dqf.sum <- function(dqf){
  n.functions <- length(dqf[,1])
  ret <- dqf
  
  for(i in 1:n.functions){
    func.sum <- sum(dqf[i,])
    for(j in 1:length(dqf[i,])){
      ret[i,j] <- dqf[i,j]/func.sum
    }
  }
  
  return(ret)
}
```

### Scale functions (global)

#### `scale.dqf.globalmax`: Given dqfs, where rows and columns are functions' x and y values, scale each function to relative to global max.

```{r}
scale.dqf.globalmax <- function(dqf){
  n.functions <- length(dqf[,1])
  globalmax <- max(dqf)
  
  ret <- dqf
  
  for(i in 1:n.functions){
    for(j in 1:length(dqf[i,])){
      ret[i,j] <- dqf[i,j]/globalmax
    }
  }
  
  return(ret)
}
```

#### `scale.dqf.globalsum`: Given dqfs, where rows and columns are functions' x and y values, scale each function to relative to global sum.

```{r}
scale.dqf.globalsum <- function(dqf){
  n.functions <- length(dqf[,1])
  globalsum <- sum(dqf)
  
  ret <- dqf
  
  for(i in 1:n.functions){
    
    for(j in 1:length(dqf[i,])){
      ret[i,j] <- dqf[i,j]/globalsum
    }
  }
  
  return(ret)
}
```































